\LARGE \bf GPU Acceleration for Fixed Complexity Sphere Decoder in Large MIMO Uplink Systems

### 2015-10-21 FCSD-CUDA project/parallel_acceleration_of_FCSD_based_on_cuda/conference draft.tex
revised version after acceptance 1
simulation result for MMSE MMSE-OSIC MMSE-AOMIC, MMSE-OMIC, RSVD, RSVD-OMIC

General Purpose Graphic Processing Units (GPGPU) computing is a common trend in industry and research in areas that need massive computations, such as computer vision\cite{fung2008using}, computer games\cite{blewitt2013applicability}, signal processing\cite{van2011accelerating}\cite{6671435}, and finance\cite{grauer2013accelerating}. Compared to a CPU, a GPU can naturally cope with computational intensive tasks, since most of the hardware resources are allocated for data processing and parallel computation, rather than caching and flow control\cite{cook2013cuda}. In recent years GPU computing has provided significant advantages in data processing speeds over CPU computations\cite{nvidia2008programming}. The Computed unified Device Architecture (CUDA)\cite{cook2013cuda}\cite{nvidia2008programming} is a programming model for GPGPU that was proposed by NIVIDIA and has become the dominant framework for GPGPU programming. It is based on popular programming languages (C, C++, Java, Python, Fortran, Direct Compute), such that the application developers in different areas do not need to learn a new programming language or get to know the hardware details. CUDA employs the Single Instruction Multiple Threads (SIMT) programming model, enables a GPU to implement parallel data processing, and has become a parallel computing platform for different applications.

In this paper, a GPU based implementation of Fixed Complexity Sphere Decoder (FCSD) is proposed, which is referred to as CUDA-FCSD. A heterogeneous programming solution for FCSD preprocessing and integrated blocked parallel path searching and decision making are presented. Using this implementation the speedup as well as Bit Error Rate (BER) performance of different sizes of MIMO systems and signal constellations are analyzed. The proposed CUDA-FCSD implementation demonstrates a great potential for large MIMO systems.

@@

In Fig. \ref{BER curve} we present Monte Carlo simulation results for bit error rate (BER) performance, using the GPU implementation of different MIMO system configurations and constellation sizes. We consider uncoded complex MIMO systems with $M$QAM, where each antenna transmits different streams of binary data generated randomly and independently. The results have been obtained by using $100000$ channel realizations with at least $500$ symbol errors accumulated. The simulation results show that FCSD-CUDA provides the same BER performance as its CPU version\cite{barbero2008fixing}.
\begin{figure}[htb]
\centering
\includegraphics*[width=0.56\textwidth]{BER_curves.eps}
\caption{BER performance for MIMO systems with various modulation schemes using the GPU implementation}
\label{BER curve}
\end{figure}
\section{Conclusions}\label{conclusion}
This paper presents a GPU implementation of the fixed complexity sphere decoder (FCSD) for large MIMO systems. In order to exploit the computational capability of GPU in the simulation of large MIMO systems, the utilization of a heterogeneous programming model, accounting for resource limitations and memory configuration must be considered. The simulation results show that a considerable acceleration of the GPU implementation of FCSD can be reached for large MIMO systems and signal constellation sizes over CPU implementation, while maintaining the same BER performance. Since the GPU and CPU can work independently, the data processing power of GPU computing is an additional advantages provided by the traditional simulation platforms. This shows the potential of GPU computing to reduce the time associated with intensive simulations of large MIMO systems.

>>>>

General Purpose Graphic Processing Units (GPGPU) computing is a common trend in industry and research in areas that need massive computations. Compared to a CPU, a GPU can naturally cope with computational intensive tasks, since most of the hardware resources are allocated for data processing and parallel computation, rather than caching and flow control\cite{cook2013cuda}. In recent years GPU computing has provided significant advantages in data processing speeds over CPU computations\cite{nvidia2008programming}. GPU computing has been applied to different research fields such as computer vision\cite{fung2008using}, computer games\cite{blewitt2013applicability}, signal processing\cite{van2011accelerating}\cite{6671435}, and finance\cite{grauer2013accelerating}. The Computed unified Device Architecture (CUDA)\cite{cook2013cuda}\cite{nvidia2008programming} is a programming model for GPGPU that was proposed by NIVIDIA and has become the dominant framework for GPGPU programming. It is based on popular programming languages (C, C++, Java, Python, Fortran, Direct Compute), such that the application developers in different areas do not need to learn a new programming language or get to know the hardware details. CUDA employs the Single Instruction Multiple Threads (SIMT) programming model, enables a GPU to implement parallel data processing, and has become a parallel computing platform for different applications.

 \cite{hongyuan2010fixed} presents an interesting implementation of Fixed Complexity Sphere Decoder (FCSD) based on GPU, however only parallel implementation of multiple decoders is considered rather than exploiting the attribute of parallelism of the algorithm itself. In this paper, a GPU based implementation of single FCSD is proposed, which is referred to as CUDA-FCSD. A heterogeneous programming solution for FCSD preprocessing and integrated blocked parallel path searching and decision making are presented. Using this implementation the speedup as well as Bit Error Rate (BER) performance of different sizes of MIMO systems and signal constellations are analyzed. The proposed CUDA-FCSD implementation demonstrates a great potential for large MIMO systems.

@@

In Fig. \ref{BER curve} we present Monte Carlo simulation results for bit error rate (BER) performance, using the GPU implementation of different MIMO system configurations and constellation sizes. We consider uncoded complex MIMO systems with $M$ QAM, where each antenna transmits different streams of binary data generated randomly and independently. The results have been obtained by using $10^{5}$ channel realizations with at least $500$ symbol errors accumulated. The simulation results show that FCSD-CUDA provides the same BER performance as its CPU version, due to the space limitation we do not present the results of CPU implementation, \cite{barbero2008fixing} provides some of the referable results.
\begin{figure}[htb]
\centering
\includegraphics*[width=0.5\textwidth, height=9.5cm]{BER_curves.eps}
\caption{BER performance for MIMO systems with various modulation schemes using the GPU implementation}
\label{BER curve}
\end{figure}
\section{Conclusions}\label{conclusion}
This paper presents a GPU implementation of the fixed complexity sphere decoder (FCSD) for large MIMO systems. In order to exploit the computational capability of GPU in the simulation of large MIMO systems, the utilization of a heterogeneous programming model, accounting for resource limitations and memory configuration must be considered. The simulation results show that significant accelerations of the GPU implementation of FCSD can be reached for large MIMO systems and signal constellation sizes over CPU implementation, while maintaining the same BER performance. Since the GPU and CPU can work independently, the data processing power of GPU computing is an additional advantages provided by the traditional simulation platforms. This shows the potential of GPU computing to reduce the time associated with intensive simulations of large MIMO systems.
